\Large \bf Uncovering Distributed System Bugs during Testing (not Production!)

### 2015-09-20 overview.tex
updates in section 3

The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines. A key capability of the \psharp runtime is that it can execute in bug-finding mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in Section~\ref{sec:bg:bugs}.

@@

\subsubsection{Handling \csharp 5.0 asynchrony}
\label{sec:psharp:async}

The Live Azure Migration system (Section~\ref{sec:cases:migration}) uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives.
An \texttt{async} procedure is allowed to perform \texttt{await} on a TPL task without blocking the current thread.
The \csharp compiler achieves this non-blocking behavior by wrapping the \emph{continuation} of the \texttt{await} statement
in a TPL task and returning it to the caller.
The TPL scheduler executes this task when the task being awaited has completed.
In principle, this asynchrony can be modeled using the \psharp primitives for machine creation and message passing.
However, in our experience, refactoring code to model in \psharp the asynchrony in async/await is difficult and unlikely
to scale for large code bases.
Therefore, our implementation automatically captures asychronously created tasks in .NET runtime as \psharp machines.

Our solution involves using a custom task scheduler whenever the bug-finding mode of the \psharp runtime is enabled, instead of the default TPL task scheduler. The \psharp task scheduler inherits from the \texttt{TaskScheduler} class, a low-level API that is responsible for enqueueing TPL tasks into threads. Our approach works as follows. We start the task of the \emph{root} \psharp machine in our custom task scheduler. As long as any child tasks spawned by the root machine task do not explicitly start in another scheduler (including the default TPL scheduler), then they will be scheduled for execution in our custom task scheduler. Our custom scheduler intercepts the call to enqueue a task, creates a special machine (only internally available to \psharp) called \texttt{TaskMachine} and \emph{wraps} the enqueued task inside this machine. The constructor of a \texttt{TaskMachine} takes as an argument a TPL task and stores it in an field. The \texttt{TaskMachine} has only a single state and when it is scheduled for execution by the \psharp systematic testing scheduler it starts executing the wrapped task and then immediately waits on its completion. This forces the lifetime of the task to become the lifetime of the machine and vice versa. Thus, when the \psharp systematic testing scheduler schedules the \texttt{TaskMachine}, it will also schedule the corresponding task. This means that when a \psharp program makes a call to an \texttt{async} method, then the task created in the backend will be automatically wrapped and scheduled.

For \texttt{await} statements, no special treatment is required because \texttt{await} statements are compiled into a continuation. The task associated with the continuation will also be wrapped and scheduled accordingly.

>>>>

The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines. A key capability of the \psharp runtime is that it can execute in \emph{bug-finding} mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in Section~\ref{sec:bg:bugs}.

@@

\textbf{Handling \csharp 5.0 asynchrony}.
The Live Azure Table Migration system (see Section~\ref{sec:cases:migration}) heavily uses the \texttt{async} and \texttt{await} \csharp 5.0 language primitives. An \texttt{async} method is able to perform \texttt{await} on a TPL task \emph{without} blocking the current thread. The \csharp compiler achieves this non-blocking behavior by wrapping the code following an \texttt{await} statement in an internal TPL \emph{task continuation}, and returning execution to the caller of the \texttt{async} method. The TPL scheduler executes this task continuation when the task being awaited has completed. In principle, this asynchrony can be modeled using the \psharp primitives for machine creation and message passing. However, in our experience, modeling async/await asynchronous code using \psharp is difficult and time consuming, and thus unlikely to scale for large code bases. To tackle this problem, we developed a solution that automatically captures asynchronously created TPL tasks and wraps them in temporary \psharp machines.

Our solution involves using a \emph{custom task scheduler}, during systematic testing with \psharp, instead of the default TPL scheduler. Assuming that the developer does not explicitly change the task scheduler (something extremely uncommon), any TPL tasks spawned in the \psharp program will be scheduled for execution in the \psharp task scheduler. These tasks will be intercepted by our custom task scheduler and wrapped inside a special, available only internally, task-machine. When the \psharp testing engine schedules a task-machine for execution, the corresponding task is unwrapped and executed. This approach works nicely with the \texttt{async} and \texttt{await} \csharp 5.0 primitives: when an \texttt{async} method is called, the asynchronous TPL task will be automatically wrapped in a task-machine and scheduled by \psharp; for \texttt{await}, no special treatment is required since the task continuation will also be automatically wrapped and scheduled accordingly.
